\documentclass[12pt,spanish]{memoir}
\usepackage{style/cleanbook}

\author{Miguel A. Dur\'{a}n-Olivencia}
\title{{\big Topology}}
\date{Brief notes - October/2020}

\begin{document}

\thispagestyle{empty}
\maketitle
\thispagestyle{empty}
\newpage

\frontmatter
\tableofcontents

\mainmatter
%\input{chapters/ch-1.tex}
\chapter{Topological spaces}

In this first chapter, we are going to study a fundamental concept which will pervade the rest of this course, the concept of \emph{topological space}. Once this concept is introduced, we will proceed with some basic topological concepts, besides studying the most important operations which allow for the construction of new topological spaces.

\section{Topologies, bases and subbases}

\subsection{Topologies}
The modern definition of topological space is based upon the fundamental concept of open set (which is axiomatically introduced). Thus, a topological space is defined by specifying the family of open sets which form it.  The formal definition is as follows.

\begin{definition}
Let $X$ be a set, and let $\T$ be a family of subsets of $X$ (i.e., $\T\subset\P X$). Then $\T$ is called a \textbf{topology} on $X$ if the following properties are fulfilled:
\begin{enumerate}
	\item $\emptyset, X\in \T$
	\item Let $\{A_i\}_{i\in I}$ be a family of subsets such that $A_i \in \T$, then $\bigcup\limits_{i\in I} A_i\in\T$
	\item Let $A, B\in \T$, then $(A\cap B)\in \T$
\end{enumerate}
\label{def:topology}
\end{definition}

Having the formal definition for \emph{topology}, it readily follows that a \textbf{topological space} is a tuple $(X, \T)$, with $X$ being a set and $\T$ a topology on $X$. The elements of $\T$ are called the \textbf{open subsets} of $X$. 

\begin{remark}
The property (\defref{def:topology}-2) could be rewritten in a much more concise by saying that, if $\F\subset\T$, then $\bigcup\F\in\T$.
Also, the third property can be readily extended to any finite amount of open sets by induction, so that if $\{A_k\}_{k=1,\dots,n}$ with $A_k\in\T$, then $\bigcap\limits_{k=1}^nA_k\in \T$.
\end{remark}

For practical purposes, we might simply use $X$ instead of $(X, \T)$, keeping in mind that $\T$ changes with $X$. 
%
The first topological concept we introduce based upon the concept of \emph{open set} is the concept of \emph{neighbourhood}.

\begin{definition}
Let $C, U\subset X$, with $(X, \T)$ being a topological space. Then, $U$ is said to be a \textbf{neighbourhood} of $C$ if $\exists A\in\T$ such that $C\subset A \subset U$. The subset $U$ is said to be a neighbourhood of a point $x\in X$ if $U$ is a neighbourhood of $\{x\}$. An open neighbourhood of $x$ is simply an open subset of $X$ containing $x$. 
\label{def:neighbourhood}
\end{definition}

The neighbourhoods of the points of a topological space determine the open sets of the topology:

\begin{theorem}
Let $(X, \T)$ be a topological space. A subset $U\subset X$ is an open subset \emph{if and only if} (iff) $U$ is neighbourhood of all its points.
\end{theorem}
\begin{proof} To prove this theorem will proceed with the demonstration of the left and right implies.

\noindent{$\boxed{\Rightarrow}$} This is trivial, given that $U\in\T$ exists and fulfils the conditions to be a neighbourhood of all its points.

\noindent{$\boxed{\Leftarrow}$} Let $U$ be a neighbourhood of all its points. Then, $\forall x\in U$, $\exists V\in \T\,:\,\{x\}\subset V\subset U$ $\Rightarrow$ $U = \bigcup\F$ with $\F=\{V\in\T\,:\,V\subset U\}\,\Then\,U$ is open, since the union of open subsets is open according to \defref{def:topology}-2.
\end{proof}

The intuitive idea to develop so far is that a neighbourhood of a point $x$ is a set $U$ which contains all the points ``around'' $x$, which also means that $U$ is open \iff this set only contains all the points surrounding every of its points.

\begin{example}
Let $X$ be a set, the \emph{trivial topology} is $\T=\{\emptyset, X\}$, and the topology $\T=\P X$ is called the $\emph{discrete topology}$.
\end{example}

The proof that both examples fulfil the conditions required by the definition \defref{def:topology} is straightforward, and left to the reader if interested. On the one hand, the trivial topology is the smallest one possible, since it has the minimal elements required to fulfil the definition \defref{def:topology}. On the other hand, the discrete topology is the largest one possible, given that all the possible subsets of $X$ are open.

\begin{definition}
The elements $x\in X$ will be called \emph{points}. Thus, a point $x\in X$ is said to be \emph{isolated} if $\{x\}$ is an open subset of $X$, i.e. $\{x\}\in \T$. 
\end{definition} 

\begin{remark} 
A trivial result is that a topological space $(X,\T)$ is \emph{discrete} (i.e., $\T=\P X$) \iff\ $\forall x\in X$, $x$ is isolated.
\end{remark}

\subsection{Bases}

\begin{definition}
Let $(X, \T)$ be a topological space, and $\B = \{B \subset X\,:\, B \in \T\}\subset\T$ a family of open sets. The family $\B$ is a \textbf{base} of $X$ \iff\, every open subset of $X$ can be expressed as the union of elements of $\B$. In other words, $\B\subset\T$ is a base of $(X,\T)$ if for every $U\in\T$ and $u\in U$, there is a $B\in \B$ such that $u\in B\subset U$.
\label{def:base}
\end{definition}

The convenience of introducing this concept becomes clear with the following two theorems:

\begin{theorem}
Let $(X, \T)$ be a topological space, and $\B$ a base of $(X, \T)$.  The following two properties are satisfied: 
\begin{enumerate}
\item $\bigcup\limits_{B\in\B} B \equiv \bigcup\B = X$
\item Let $A, B\in \B$ and $x\in A\cap B$, then there exists $C\in \B$ such that $x\in C\subset (A\cap B)$
\end{enumerate}
\end{theorem}

\begin{proof} 
This proof is straightforward by using \defref{def:topology} and \defref{def:base}:
\begin{enumerate}
\item \noindent{$\boxed{\Rightarrow}$}  $\bigcup\B \subset X$ is trivial.

\noindent{$\boxed{\Leftarrow}$} Given that $X\in \T$ (hence it is an open subset) $\Then$ $X$ can be expressed as the union of a family $\F\subset \B$ $\Then$ $X \subset \bigcup \F\subset \bigcup \B$.

\item Let $A, B\in \B$ $\Then$ $(A\cap B)\in \T$ $\Then$ according to \defref{def:base}, $\exists C\in \B$ such that $x\in C\subset(A\cap B)$.
\end{enumerate}
With this, the theorem is proved to hold true. 
\end{proof}

Interestingly, the two properties of the previous theorem are sufficient for a set family to be a base of a topology:

\begin{theorem}
Let $X$ be a set and $\B\subset \P X$ a family of subsets of $X$ such that:
\begin{enumerate}
\item $\bigcup \B = X$
\item If $A, B\in \B$ and $x\in (A\cap B)$, there exists a $C\in \B$ such that $x\in C\subset (A\cap B)$
\end{enumerate}
Therefore, there exists a unique topology $\T$ on $X$ with $\B$ being a base of it, and such topology is the smallest (w.r.t.\footnote{with respect to} the inclusion) for which the elements of $\B$ are open.
\label{th:1-3}
\end{theorem}

\begin{proof}
To proceed with the proof, we will first proof that there exist a topology, and later that this is unique.
Let us define $\T = \{\bigcup \F\,:\,\F\subset \B\}$, and now proceed with the proof that $\T$ is a topology.
\begin{enumerate}
\item Then, by using the first point $\T =\bigcup \B= X$ $\Then$ $X\in \T$. And, it is trivial that $\emptyset = \bigcup \emptyset \in \T$.
\item Let $\F\subset \T$, and let $\F_A=\{B\in \B\,:\, B\subset A\}$ for each $A\in\F$, so that $A=\bigcup\F_A$. Let $\F^*=\bigcup\limits_{\F_A\in\F}\F_A\subset\B$. Then,  $\bigcup \F =\bigcup\F^*\in\T$.
\item Let $A, B\in \T$, and $\F = \{B\in\B\,:\, B \subset(A\cap B)\}$. Let us prove that $A\cap\B = \bigcup\F$:
\begin{itemize}
\item[\boxed{\supseteq}] This is a trivial proof: $\forall x\in\bigcup\F$ $\Then$ $x\in\setdef{B\in\B}{B\subset(A\cap B)}\ \Then\ x\in (A\cap B)$ $\Then$ $\bigcup\F \subset (A\cap B)$.
\item[\boxed{\subseteq}] Let $x\in (A\cap B)$, according to $(\emph{2.})$, $\exists C\in \B$ such that $x\in C\subset (A\cap B)$ $\Then$ $x\in C\subset \bigcup\F$ $\Then$ $(A\cap B)\subset \bigcup\F$. 
\end{itemize}
Finally, $\bigcup \F\in \T$ $\Then$ $(A\cap B) \in \T$.
\end{enumerate}
Therefore, $\T$ is a topology on $X$. To demonstrate the final part of the theorem, consider $\T'$ a topology on $X$ with $\B$ as base. By definition, $\T\subset \T'$. Therefore, $\T$ is unique and minimal.
\end{proof}

Thus, to define a topology on a set $X$ one only needs to determine a family $\B$ with the properties just described.
%
In what follows, we introduce another related concept which will be convenient later.
%
\begin{definition}
Let $(X, \T)$ be a topological space, and $x\in X$. A \textbf{base of (open) neighbourhoods} of $x$ is a family $\B\subset\T$ of open neighbourhoods of $x$ such that, for every $x\in U\in \T$, $\exists B\in\B$ such that $x\in B\subset U$.
\label{def:base-open-neighbourhoods}
\end{definition}

It is clear that if $(X,\T)$ is a topological space, and we have a family $\{\B_x\}_{x\in X}$ so that $\B_x$ is a base of neighbourhoods of $x$, then $\B=\bigcup\limits_{x\in X}\B_x$ is a base on (X, $\T$). Reciprocally, let $\B$ be a base on $(X, \T)$, then $\B_x = \setdef{A\in\B}{x\in A}$ is a base of neighbourhoods of $x$.

\subsection{Subbases}

In this section we will introduce a concept similar to \emph{base}, which will also be useful when defining topologies.

\begin{definition}
Let $(X, \T)$ be a topological space. A family $\S\subset \T$ is a \textbf{subbase} of $(X, \T)$ \iff\ the family of finite intersections of elements of $\S$,  i.e. $\F = \setdef{\bigcap\limits_{i=1}^nS_k}{S_k\subset \S}$\footnote{Here we accept that $\bigcap \emptyset = X$.}, is a base of $(X, \T)$.
\label{def:subbase}
\end{definition}

The difference between subbases and bases is that the family of subsets does not have to satisfy any specific condition to be a subbase:

\begin{theorem}
Let $X$ be a set, and $\S\subset \P X$ a family of subsets of $X$. There exists a unique topology on $X$ with $\S$ as subbase, and it is smallest (w.r.t. the inclusion) for which the elements of $\S$ are open.
\end{theorem}

\begin{proof}
The proof 
\end{proof}

\end{document}
